Beating Simple Backtracking Search in Subgraph Algorithms\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant numbers EP/026842/1 and EP/M508056/1].}

### 2018-11-22 paper/paper.tex
More experiments

\begin{abstract}
    The current state of the art in subgraph isomorphism solving involves using degree as a
    value-ordering heuristic to direct backtracking search. Such a search makes a heavy commitment
    to the first branching choice, which is often incorrect. To mitigate this, we introduce and
    evaluate a new approach, which we call ``solution-biased search''. By combining a
    slightly-random value-ordering heuristic, rapid restarts, and nogood recording, we achieve two
    orders of magnitude faster performance on satisfiable instances, whilst not affecting
    performance on unsatisfiable instances. This search can also be parallelised in a very simple
    but effective way, giving us a further speedup of over thirty from thirty-six cores, across both
    satisfiable and unsatisfiable instances.  Finally, we show that solution-biased search is also
    suitable for optimisation problems, by using it to improve two maximum common induced subgraph
    algorithms.
\end{abstract}

@@

The subgraph isomorphism problem is to decide whether a copy of a small ``pattern'' graph occurs
inside a larger ``target'' graph. Although the problem is \NP-complete, by combining design techniques from
artificial intelligence with careful algorithm engineering, modern subgraph isomorphism solvers can
often produce exact solutions quickly even on graphs with thousands of vertices. The current single
strongest subgraph isomorphism solver uses ``highest degree first'' as a value-ordering heuristic to
direct a constraint programming style search
\cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:journals/jair/McCreeshPST18}. This
heuristic is much better than branching randomly, but is still far from perfect. We therefore
investigate different techniques for offsetting mistakes in value-ordering decisions. We first try
discrepancy searches, which turn out to be much too expensive to be a practical solution. We then
look at using random value-ordering, restarts, and nogood recording, which together give better
aggregate performance at the expense of much worse behaviour on many instances. Next, we introduce a
new way of adapting the existing value-ordering heuristic to introduce just a small amount of
randomness.  This lets us make a state-of-the-art subgraph algorithm perform much better
on a large number of satisfiable instances, whilst performing worse only rarely on satisfiable
instances, and never on unsatisfiable instances.

Finally, we show that this strategy is also effective in an optimisation setting, producing benefits
in two maximum common induced subgraph algorithms. This combination of techniques gives us, for the
first time, a practical alternative to backtracking search for constraint programming style
algorithms where we have a strong value-ordering heuristic and where we care both about satisfiable
and unsatisfiable instances. ?? Parallel

@@

Our experiments are performed on the EPCC Cirrus HPC facility, on systems with dual Intel Xeon
E5-2695 v4 CPUs and 256GBytes RAM, running Centos 7.3.1611. ?? C++ native threads, MPI. We use GCC
7.2.0 as the compiler. ?? Steady clock times. We use a deterministic pseudo-random number generator
for reproducibility.

@@

We begin by verifying that the Glasgow Subgraph Solver is the state of the art. In the first and
third plots \cref{figure:others} we show the cumulative number of instances solved over time for the
non-induced and induced problems respectively. We compare the Glasgow Subgraph Solver (both
unmodified, and with the modifications described in the remainder of this paper), the PathLAD
variation of the LAD algorithm \cite{DBLP:journals/ai/Solnon10,DBLP:conf/lion/KotthoffMS16}, VF2
\cite{DBLP:journals/pami/CordellaFSV04}, RI \cite{DBLP:journals/bmcbi/BonniciGPSF13}, and VF3
\cite{DBLP:conf/gbrpr/CarlettiFSV17} (which only supports the induced problem), in each case using
the original implementation provided by the algorithm's authors. For each algorithm, the $y$ value
gives the cumulative number of instances which (individually) can be solved in no more than $x$
milliseconds.  The vertical distance between two lines therefore shows how many more instances can
be solved by one solver than another, if every instance is run separately with the chosen $x$
timeout. The horizontal distance shows how many times longer the per-instance timeout would need to
be to allow the rightmost algorithm to succeed on $y$ out of the 14,621 instances (bearing in mind
that the two sets of $y$ instances could be different).  The plots show that our starting point
comfortably beats PathLAD, VF2, VF3 and RI, except for very low choices of timeout.

@@

?? Also do LDS?

@@

\begin{figure}[p]
    \centering
    \includegraphics*{gen-graph-parallel.pdf}

@@

    \includegraphics*{gen-graph-scatter-par-repeat.pdf}
    \hfill
    \includegraphics*{gen-graph-scatter-parconsttick-repeat.pdf}

@@

\begin{figure}[tb]
    \includegraphics*{gen-graph-scatter-parconst-nodes.pdf}
    \hfill
    \includegraphics*{gen-graph-scatter-parconsttick-nodes.pdf}

    \caption{Parallel nodes}\label{figure:parallelnodes}
\end{figure}

>>>>

\begin{abstract}
    The current state of the art in subgraph isomorphism solving involves using degree as a
    value-ordering heuristic to direct backtracking search. Such a search makes a heavy commitment
    to the first branching choice, which is often incorrect. To mitigate this, we introduce and
    evaluate a new approach, which we call ``solution-biased search''. By combining a
    slightly-random value-ordering heuristic, rapid restarts, and nogood recording, we achieve two
    orders of magnitude faster performance on satisfiable instances, whilst not affecting
    performance on unsatisfiable instances. This search can also be parallelised in a very simple
    but effective way, giving us a further speedup of over ?? from thirty-six cores, across both
    satisfiable and unsatisfiable instances.  Finally, we show that solution-biased search is also
    suitable for optimisation problems, by using it to improve two maximum common induced subgraph
    algorithms.
\end{abstract}

@@

The subgraph isomorphism problem is to decide whether a copy of a small ``pattern'' graph occurs
inside a larger ``target'' graph. Although the problem is \NP-complete, by combining design
techniques from artificial intelligence with careful algorithm engineering, modern subgraph
isomorphism solvers can often produce exact solutions quickly even on graphs with thousands of
vertices. The current single strongest subgraph isomorphism solver uses ``highest degree first'' as
a value-ordering heuristic to direct a constraint programming style search
\cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:journals/jair/McCreeshPST18}. This
heuristic is much better than branching randomly, but is still far from perfect. To offset mistakes
made by this heuristic, this paper proposes a new perspective on value-ordering: rather than
defining a search order, we use degree to direct what \emph{proportion} of the search effort should
be spent in each subproblem. By combining rapid restarts and nogood recording, and introducing a
small amount of randomness into the value-ordering heuristic, we make a state-of-the-art subgraph
algorithm perform two orders of magnitude better on a large number of satisfiable instances, whilst
performing worse only rarely on satisfiable instances, and never on unsatisfiable instances.

This new form of search can also be parallelised. By running many threads with different random
seeds but the same restart schedule, and sharing nogoods only following restarts, we can achieve
speedups of ??, with a \emph{much} simpler implementation than conventional work-stealing.  Finally,
we show that this strategy is also effective in an optimisation setting, producing benefits in two
maximum common induced subgraph algorithms. This combination of techniques gives us, for the first
time, a practical alternative to backtracking search for constraint programming style algorithms
where we have a strong value-ordering heuristic and where we care both about satisfiable and
unsatisfiable instances.

@@

Our experiments are performed on the EPCC Cirrus HPC facility, on systems with dual Intel Xeon
E5-2695 v4 CPUs and 256GBytes RAM, running Centos 7.3.1611. We use GCC 7.2.0 as the compiler. For
paralellism, we use C++ native threads, and for distributed parallelism we also use the SGI MPT
implementation of MPI. All timing measurements are steady-clock, and we use a deterministic
pseudo-random number generator for reproducibility.

@@

We begin with a set of baseline performance measurements. In the first and
third plots \cref{figure:others} we show the cumulative number of instances solved over time for the
non-induced and induced problems respectively. We compare the Glasgow Subgraph Solver using
depth-first search (DFS) and with the modifications described in the remainder of this paper
(solution-biased search, SBS), the PathLAD variation of the LAD algorithm
\cite{DBLP:journals/ai/Solnon10,DBLP:conf/lion/KotthoffMS16}, VF2
\cite{DBLP:journals/pami/CordellaFSV04}, RI \cite{DBLP:journals/bmcbi/BonniciGPSF13}, and VF3
\cite{DBLP:conf/gbrpr/CarlettiFSV17} (which only supports the induced problem), in each case using
the original implementation provided by the algorithm's authors. For each algorithm, the $y$ value
gives the cumulative number of instances which (individually) can be solved in no more than $x$
milliseconds.  The vertical distance between two lines therefore shows how many more instances can
be solved by one solver than another, if every instance is run separately with the chosen $x$
timeout. The horizontal distance shows how many times longer the per-instance timeout would need to
be to allow the rightmost algorithm to succeed on $y$ out of the 14,621 instances (bearing in mind
that the two sets of $y$ instances could be different).  The plots show that our starting point
comfortably beats PathLAD, VF2, VF3 and RI, except for very low choices of timeout.

@@

@@

\begin{figure}[tb]
    \centering
    \includegraphics*{gen-graph-parallel.pdf}

@@

@@

% \begin{figure}[tb]
%     \includegraphics*{gen-graph-scatter-par-repeat.pdf}
%     \hfill
%     \includegraphics*{gen-graph-scatter-parconsttick-repeat.pdf}
%
%     \includegraphics*{gen-graph-scatter-parconst-nodes.pdf}
%     \hfill
%     \includegraphics*{gen-graph-scatter-parconsttick-nodes.pdf}
%
%     \caption{Parallel nodes}\label{figure:parallelnodes}
% \end{figure}
